---
title: "Prosper Loan Data Analysis by Yuxun Ling"
output: html_document
---
##Loading Dataset and Libraries
```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(GGally)
library(scales)
```

```{r echo=FALSE, Load_The_Data}
#Load the data
data_o <- read.csv('prosperLoanData.csv', sep = ',')
```

```{r echo=FALSE}
str(data_o)
```
##Some Basic Overviews on The Dateset
```{r echo=FALSE}
#find all categorical variables
data_sub_cat <- data_o[sapply(data_o,is.factor)]
names(data_sub_cat)
```
 


```{r}
#identify levels of factors
levels(data_o$CreditGrade)
levels(data_o$LoanStatus)
levels(data_o$ProsperRating..Alpha.)
levels(data_o$BorrowerState)
levels(data_o$Occupation)
levels(data_o$EmploymentStatus)
levels(data_o$IncomeRange)
```

```{r echo=FALSE}
#summarise the data
summary(data_o)
```

The summary indicates that ListingKey, ListingCreationDate, LoanKey, and DateCreditPulled are related
  
ProsperRating and ProsperScore related to EstimatedEffectiveYield, EstimatedLoss, and EstimatedRetuen. (because they have same amount of NA's)
  
Over 2/3 of the total observations do not have CreditGrade
  
Some variables related to delinquencies have skewed obervations.

##Univariante Graphs and Basic Analysis
```{r warning=FALSE, echo=FALSE, message=FALSE}
#Univariate Graphs
qplot(data = subset(data_o,CreditGrade!=''),x=CreditGrade)
qplot(data=data_o,x=Term)
```

Only Three Possible Terms are allowed. (One year, Three years, and five years)
 
```{r warning=FALSE, echo=FALSE, message=FALSE}
qplot(data=data_o,x=BorrowerAPR)
qplot(data=data_o,x=BorrowerRate)
qplot(data=data_o,x=LenderYield)
qplot(data=data_o,x=EstimatedEffectiveYield)
qplot(data=data_o,x=EstimatedLoss)
```

The above five graphs are quite similar because BorrowerAPR, EstimatedEffectiveYield and LenderYield are BorrowerRate minusing some comparatively small percentages or constants (e.g. servicing fee), and EstimatedLoss should be a percentage of BorrowerRate plusing some constant.
 
```{r warning=FALSE, echo=FALSE, message=FALSE}
cor.test(data_o$BorrowerAPR,data_o$BorrowerRate,method = 'pearson')
cor.test(data_o$LenderYield,data_o$BorrowerRate,method = 'pearson')
cor.test(data_o$EstimatedEffectiveYield,data_o$BorrowerRate, method = 'pearson')
cor.test(data_o$EstimatedLoss,data_o$BorrowerRate,method = 'pearson')
```

The pearson correlation test also suggests strong correlation between BorrowerRate and other four variables.
  
The correlation between EstimateedEffectiveYield and BorrowerRate is 0.5% less than 0.9 may due to the involving of late fee and charge-off loss.
 
```{r warning=FALSE, echo=FALSE, message=FALSE}
qplot(data = data_o,x=EstimatedReturn)
qqnorm(with(subset(data_o,!is.na(EstimatedReturn)),EstimatedReturn-mean(EstimatedReturn)))
```

EstimatedReturn is not symmetrically distributed around mean.
 
```{r warning=FALSE, echo=FALSE, message=FALSE}
qplot(data=data_o,x=EstimatedReturn,bins=500)
data_o.sum_by_EstRet <- subset(data_o,!is.na(EstimatedReturn)) %>%
  group_by(EstimatedReturn) %>%
  summarise(n=n()) %>%
  arrange(desc(n))
head(data_o.sum_by_EstRet)
```

A closer look indicates that some specific values (e.g. 0.1246) have extremely high count.
 
```{r warning=FALSE, echo=FALSE, message=FALSE}
data_o.sum_by_BorRat <- subset(data_o,!is.na(BorrowerRate)) %>%
  group_by(BorrowerRate) %>%
  summarise(n=n()) %>%
  arrange(desc(n))
head(data_o.sum_by_BorRat)
```

Counting the frequency of each value of BorrowerRate. One possibile cause of high-count EstimateReturn may be a high count of BorrowerRate at 0.3177.
 
```{r warning=FALSE, echo=FALSE, message=FALSE}
# qplot(data=data_o,x=ProsperRating..numeric.)
# qplot(data=data_o,x=ProsperScore)
qplot(data=data_o,x=EmploymentStatusDuration)
```
```{r warning=FALSE, echo=FALSE, message=FALSE}
ggplot(data = data_o,aes(x=EmploymentStatusDuration))+
  geom_histogram(binwidth = 3)+
  xlim(c(0,600))
```

Notice that in average, people with longer employment duration are less likely to borrow.
 
```{r warning=FALSE, echo=FALSE, message=FALSE}
qplot(data=data_o,x=CreditScoreRangeLower)
qplot(data=data_o,x=CreditScoreRangeUpper)
with(data_o,cor.test(CreditScoreRangeLower,CreditScoreRangeUpper))
```
```{r warning=FALSE, echo=FALSE, message=FALSE}
# qplot(data=data_o,x=CurrentCreditLines)
# qplot(data=data_o,x=TotalCreditLinespast7years)
# qplot(data=data_o,x=OpenRevolvingAccounts)
qplot(data=data_o,x=OpenRevolvingMonthlyPayment)
qplot(data=data_o,x=InquiriesLast6Months)
qplot(data=data_o,x=TotalInquiries)
qplot(data=data_o,x=CurrentDelinquencies)
qplot(data=data_o,x=AmountDelinquent)
qplot(data=data_o,x=DelinquenciesLast7Years)
qplot(data=data_o,x=PublicRecordsLast10Years)
qplot(data=data_o,x=PublicRecordsLast12Months)
qplot(data=data_o,x=RevolvingCreditBalance)
```

Above nine graphs have extremely high count near zero, and extremely high value with low counts, and thus may not be useful on their own.
 
```{r warning=FALSE, echo=FALSE, message=FALSE}
qplot(data=data_o,x=BankcardUtilization)
ggplot(data=data_o,aes(x=BankcardUtilization))+geom_histogram(binwidth = 0.01)+xlim(0,quantile(data_o$BankcardUtilization,0.99,na.rm = TRUE))+ylim(0,2500)
```

about 1% out of all borrower have BantcardUtilization greater than 1.
 
```{r warning=FALSE, echo=FALSE, message=FALSE}
qplot(data=data_o,x=AvailableBankcardCredit)
ggplot(data = data_o,aes(x=AvailableBankcardCredit))+geom_histogram()+xlim(0,quantile(data_o$AvailableBankcardCredit,0.999,na.rm = TRUE))+scale_y_log10()
```

Notice that count of AvailableBankCredit decreases almost exponentially as the value increases.
 
```{r warning=FALSE, echo=FALSE, message=FALSE}
# qplot(data=data_o,x=TotalTrades)
# qplot(data=data_o,x=TradesNeverDelinquent..percentage.)
# qplot(data=data_o,x=TradesOpenedLast6Months)
# qplot(data=data_o,x=DebtToIncomeRatio)
# qplot(data=data_o,x=StatedMonthlyIncome)
# qplot(data=data_o,x=TotalProsperLoans)
qplot(data=data_o,x=TotalProsperPaymentsBilled)
qplot(data=data_o,x=OnTimeProsperPayments)
```

Above two graphs are quite similar
 
```{r warning=FALSE, echo=FALSE, message=FALSE}
with(data_o,cor.test(TotalProsperPaymentsBilled,OnTimeProsperPayments))
```

High correlation indicates that fewer people pay late.
 
```{r warning=FALSE, echo=FALSE, message=FALSE}
qplot(data=data_o,x=ProsperPaymentsLessThanOneMonthLate)+ylim(0,100)
# qplot(data=data_o,x=ProsperPaymentsOneMonthPlusLate)
# qplot(data=data_o,x=ProsperPrincipalBorrowed)
# qplot(data=data_o,x=ProsperPrincipalOutstanding)
ggplot(data=data_o,aes(x=ScorexChangeAtTimeOfListing))+geom_histogram(binwidth = 1)
```

There are some extremely high counts for some of the scores.
 
```{r warning=FALSE, echo=FALSE, message=FALSE}
data_o.sum_by_Scorex <- subset(data_o,!is.na(ScorexChangeAtTimeOfListing)) %>%
  group_by(ScorexChangeAtTimeOfListing) %>%
  summarise(n=n()) %>%
  arrange(desc(n))
head(data_o.sum_by_Scorex)
```
 
0, +/-40, and +/-80 are the outliers.

```{r warning=FALSE, echo=FALSE, message=FALSE}
# qplot(data=data_o,x=LoanCurrentDaysDelinquent)
# qplot(data=data_o,x=LoanFirstDefaultedCycleNumber)
# qplot(data=data_o,x=LoanMonthsSinceOrigination)
# qplot(data=data_o,x=LoanOriginalAmount)
# qplot(data=data_o,x=MonthlyLoanPayment)
# qplot(data=data_o,x=LP_CustomerPayments)
# qplot(data=data_o,x=LP_CustomerPrincipalPayments)
# qplot(data=data_o,x=LP_InterestandFees)
# qplot(data=data_o,x=LP_ServiceFees)
# qplot(data=data_o,x=LP_CollectionFees)
# qplot(data=data_o,x=LP_GrossPrincipalLoss)
# qplot(data=data_o,x=LP_NetPrincipalLoss)
# qplot(data=data_o,x=LP_NonPrincipalRecoverypayments)
# qplot(data=data_o,x=PercentFunded)
ggplot(data=data_o,aes(x=Recommendations))+geom_histogram(binwidth = 1)+scale_y_log10()
data_o.sum_by_Recom <- subset(data_o,!is.na(Recommendations)) %>%
  group_by(Recommendations) %>%
  summarise(score=mean(ProsperScore,na.rm=TRUE),
            n=n()) %>%
  arrange(desc(n))
data_o.sum_by_Recom
with(data_o,cor.test(ProsperScore,Recommendations))
# qplot(data=data_o,x=InvestmentFromFriendsCount)
# qplot(data=data_o,x=InvestmentFromFriendsAmount)
# qplot(data=data_o,x=Investors)
```

Some loans have comparatively high recommendations while most loans have no recommendation. Although by summarising, loans with more recommendations tend to have higher Prosper Scores, but corelation test tells us nothing. This may due to dominantly high count of zero recommendations. 

##More Analysis

Firstly I will investigate the relationship between the nine similar distributed variables.

```{r warning=FALSE, echo=FALSE, message=FALSE}
set.seed(436346)
data_samp<-data_o[sample(1:nrow(data_o),10000),c('OpenRevolvingMonthlyPayment','InquiriesLast6Months','TotalInquiries','CurrentDelinquencies','AmountDelinquent','DelinquenciesLast7Years','PublicRecordsLast10Years','PublicRecordsLast12Months','RevolvingCreditBalance')]
ggpairs(data_samp,axisLabels='internal')
# unable to change font size of axis labels for ggpairs.
```

The graph indicates that there is no strong relationship between any pairs of them. Similarity in graphs may simply caused by large count of zeros.
 
Also, notice that there are several different scores in the dataset: CreditGrade, ProsperRating, ProsperScore, ScorexChangeAtTimeOfListing.The four scores should agree with each other.
 
```{r warning=FALSE, echo=FALSE, message=FALSE}
plot1<-ggplot(data=subset(data_o,data_o$CreditGrade!=''),aes(x=CreditGrade))+geom_bar()
plot2<-ggplot(data=subset(data_o,!is.na(data_o$ProsperScore)),aes(x=ProsperScore))+geom_bar()
plot3<-ggplot(data=subset(data_o,!is.na(data_o$ProsperRating..numeric.)),aes(x=ProsperRating..numeric.))+geom_bar()
plot4<-ggplot(data=subset(data_o,!is.na(data_o$ScorexChangeAtTimeOfListing)),aes(x=ScorexChangeAtTimeOfListing))+geom_bar()
grid.arrange(plot1,plot2,plot3,plot4,ncol=2)
```

However, Graphs of the four scores are quite different.

A question then arises: What are these four scores?

According to [the discription document](https://docs.google.com/spreadsheets/d/1gDyi_L4UvIrLTEC6Wri5nbaMmkGmLQBk-Yx3z0XDEtI/edit#gid=0), CreditGrade was the credit rating before 2009, and ProsperScore/ProsperRating was the credit rating after July 2009. The ScorexChangeAtTimeOfListing is the change in score between each loan of a specific borrower. 

```{r warning=FALSE, echo=FALSE, message=FALSE}
ggplot(aes(x=ProsperScore,y=ProsperRating..numeric.),data=data_o)+geom_point(position = 'jitter',alpha=1/50)
```

According to the graph, ProsperScore and ProsperRating agree with each other in trend, but disagree with each other in many profiles. Therefore a question arises: which score is more reliable?

```{r warning=FALSE, echo=FALSE, message=FALSE}
#create new a variable to classify profiles by year
data_o$YearCreditPulled<-as.numeric(format(as.POSIXct(data_o$DateCreditPulled),'%Y'))
```

```{r warning=FALSE, echo=FALSE, message=FALSE}
#create a subset that only contains information about members that appear twice or more.
data_o_after2009<-subset(data_o,YearCreditPulled >= 2009)

Dup_Member.member <- sort(unique(subset(data_o_after2009,!is.na(ProsperScore) & !is.na(ProsperRating..numeric.) & duplicated(MemberKey))$MemberKey))

Dup_Member.data<-subset(data_o_after2009,is.element(MemberKey,Dup_Member.member))

Dup_Member.data_earlier<-with(Dup_Member.data,aggregate(list(year=YearCreditPulled,score=ProsperScore,rating=ProsperRating..numeric.),list(id=MemberKey),min))

Dup_Member.data_later<-with(Dup_Member.data,aggregate(list(year=YearCreditPulled,delinq=AmountDelinquent),list(id=MemberKey),max))

Dup_Member.data_earlier<-Dup_Member.data_earlier[with(Dup_Member.data_earlier,order(id)),]

Dup_Member.data_later<-Dup_Member.data_later[with(Dup_Member.data_later,order(id)),]

Dup_Member_data<-cbind(Dup_Member.data_later['year']-Dup_Member.data_earlier['year'],Dup_Member.data_earlier[,c(1,3,4)],Dup_Member.data_later['delinq'])

Dup_Member_data$score<-factor(Dup_Member_data$score)
Dup_Member_data$rating<-factor(Dup_Member_data$rating)
```

```{r warning=FALSE, echo=FALSE, message=FALSE}
with(subset(Dup_Member_data,year!=0),aggregate(list(avg_delinq=delinq),list(score=score),mean))
with(subset(Dup_Member_data,year!=0),aggregate(list(avg_delinq=delinq),list(rating=rating),mean))
```

Both the Score and the Rating are reliable in the sense that people with highest scores/ratings have least average delinquencies in the 'future', and people with lowest scores/ratings have most average delinquencies in the 'future'.

However, average delinquency of people with higher ratings is lower than that of people with lower ratings, while average delinquencies of people with some scores (<2,3,4>,<5,6,7>) are approximately the same.

Therefore, the rating might be a more appropriate measure of people's credit.

The next question is: what kinds of profiles tend to have a higher rating?

The following variables may relate to the ProsperRating:

EstimatedReturn, ListingCategory, Occupation, EmploymentStatusDuration, IsBorrowerHomeowner, GroupKey, DelinquenciesLast7Years, PublicRecordsLast12Months, IncomeVerifiable, OnTimeProsperPayments, ProsperPaymentsOneMonthPlusLate, Recommendations

Effect of them will be investigated separately.
 
```{r warning=FALSE, echo=FALSE, message=FALSE}
ggplot(data=data_o,aes(y=EstimatedReturn,x=factor(ProsperRating..numeric.)))+geom_boxplot()

with(subset(data_o,!is.na(ProsperRating..numeric.)),cor.test(ProsperRating..numeric.,ListingCategory..numeric.))

ggplot(data=data_o,aes(x=Occupation,y=ProsperRating..numeric.))+geom_boxplot()+stat_summary(fun.y='mean')

with(subset(data_o,!is.na(ProsperRating..numeric.)&!is.na(EmploymentStatusDuration)),cor.test(ProsperRating..numeric.,EmploymentStatusDuration))
with(subset(data_o,!is.na(ProsperRating..numeric.)&!is.na(EmploymentStatusDuration)),aggregate(list(avg_employment_dur=EmploymentStatusDuration),list(rating=ProsperRating..numeric.),mean))

wilcox.test(subset(data_o,!is.na(ProsperRating..numeric.)&IsBorrowerHomeowner=='True')$ProsperRating..numeric.,subset(data_o,!is.na(ProsperRating..numeric.)&IsBorrowerHomeowner=='False')$ProsperRating..numeric.,alternative='greater')

subset(data_o,!is.na(ProsperRating..numeric.)&GroupKey!='') %>%
  group_by(GroupKey) %>%
  summarise(avg_rating=mean(ProsperRating..numeric.),
            n=n()) %>%
  arrange(desc(avg_rating)) %>%
  filter(n>=10) %>%
  summary()

ggplot(aes(x=factor(ProsperRating..numeric.),y=DelinquenciesLast7Years),data=subset(data_o,!is.na(ProsperRating..numeric.)))+geom_boxplot()+scale_y_log10()
with(subset(data_o,!is.na(ProsperRating..numeric.)),cor.test(ProsperRating..numeric.,DelinquenciesLast7Years))

wilcox.test(subset(data_o,!is.na(ProsperRating..numeric.)&PublicRecordsLast12Months==0)$ProsperRating..numeric.,subset(data_o,!is.na(ProsperRating..numeric.)&PublicRecordsLast12Months!=0)$ProsperRating..numeric.,alternative = 'greater')

wilcox.test(subset(data_o,!is.na(ProsperRating..numeric.)&IncomeVerifiable=='True')$ProsperRating..numeric. ,subset(data_o,!is.na(ProsperRating..numeric.)&IncomeVerifiable=='False')$ProsperRating..numeric. ,alternative = 'greater')

ggplot(aes(x=factor(ProsperRating..numeric.),y=OnTimeProsperPayments),data=subset(data_o,!is.na(ProsperRating..numeric.)))+geom_boxplot()+scale_y_log10()
with(subset(data_o,!is.na(ProsperRating..numeric.)),cor.test(ProsperRating..numeric.,OnTimeProsperPayments))

ggplot(aes(x=factor(ProsperRating..numeric.),y=ProsperPaymentsOneMonthPlusLate),data=subset(data_o,!is.na(ProsperRating..numeric.)))+geom_boxplot()+scale_y_log10()
with(subset(data_o,!is.na(ProsperRating..numeric.)),cor.test(ProsperRating..numeric.,ProsperPaymentsOneMonthPlusLate))

wilcox.test(subset(data_o,!is.na(ProsperRating..numeric.)&Recommendations!=0)$ProsperRating..numeric.,subset(data_o,!is.na(ProsperRating..numeric.)&Recommendations==0)$ProsperRating..numeric.,alternative = 'greater')
```

EstimatedReturn, IncomeVerifiable and IsBorrowerHomeowner are directly related to the rating;

Some Occupations have different mean and variance than others, but may due to the small sample size;

Mean Employement Duration of all users are approximately the same. although those who are rated 1, 2, and 3 have approximately 10% lower values than those who are rated 4, 5, and 6, those who are rated 7 have average employment duartion between the two groups;


GroupKey may have correlation with average rating, but there are just too many group keys, so it is difficult to make a conclusion about it;

DelinquentLast7Years has no strong correlation with ratings, but those who are rated 7 have significantly lower deliquencies;

Those who have PublicRecordsLast12Months equal to zero have higher average ratings by wilcox test;

ProsperPaymentsOneMonthPlusLate of those who are rated 5, 6, and 7 have decreasing means on payments that are paid more than one month late.

##Final Plots
```{r echo=FALSE, Final_Plot_1}
#Final Plot 1
ggplot(data=subset(data_o,InvestmentFromFriendsAmount!=0),aes(x=DebtToIncomeRatio ,y=InvestmentFromFriendsAmount/LoanOriginalAmount))+
  geom_point()+
  stat_sum(aes(color=factor(ListingCategory..numeric.),,size=StatedMonthlyIncome))+
  xlim(0,quantile(data_o$DebtToIncomeRatio,0.99,na.rm = TRUE))+
  ggtitle('Debt-to-Income Ratio against Investment from Friend to Total Original Loan Ratio (colored by listing category)')

ggplot(data=subset(data_o,(InvestmentFromFriendsAmount!=0 & !(ListingCategory..numeric. %in% c(0,1,4,6,7))) & IncomeVerifiable=='True'),aes(x=DebtToIncomeRatio ,y=InvestmentFromFriendsAmount/LoanOriginalAmount))+
  geom_point(alpha=0)+
  stat_sum(aes(color=factor(ListingCategory..numeric.),size=StatedMonthlyIncome))+
  scale_x_continuous(breaks=pretty_breaks(n=100))+
  scale_y_continuous(breaks=pretty_breaks(n=10))+
  xlim(0,quantile(data_o$DebtToIncomeRatio,0.99,na.rm=TRUE))+
  scale_radius(trans='sqrt',range=c(1,5))+
  ggtitle('Debt-to-Income Ratio against Investment from Friend to Total Original Loan Ratio (colored by listing filtered category)')
```

Obviously, people with heigher debt-to-income ratio have lower percentage of friend investment. Most of those who have 100% friend investment belong to category- 0 (Not Available), 1 (Debt Consolidation), 4 (Personal Loan), and 6 (Auto), which are all non-informative. By filtering out those obscure categories, I got the second graph.

Only those who are in the categories: 2 (Home Improvement), 3 (Business), 5 (Student Use),and 13 (Household expenses). Actually most loans with non-zero friend investment are for home improvement and business.

Those who applied for a loan while having a large income are very likely to be in category 2 (Home Improvement)

```{r echo=FALSE, Final_Plot_2}
#Final Plot 2
ggplot(data_o,aes(StatedMonthlyIncome,TradesNeverDelinquent..percentage.,color=IncomeVerifiable ,size=TotalTrades))+geom_point()+xlim(0,quantile(data_o$StatedMonthlyIncome,0.95))

```


